---
title: "290_data"
output: html_document
---

## Required Packages

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
```

---

## Data

Catch data in lbs. was collated for marine areas 10, 11, and 13 from Washington Department of Fish and Wildlife catch records.

## Raw data

Ending numerals in variable name represent marine area.

# Marine areas 11 & 13 catch data
```{r catch one, cache=TRUE}
DC_catch.10 <- read.csv("../Data/2007.2017_DC_catch.10.csv", header = TRUE)
```

```{r table 10, echo=FALSE, cache=TRUE}
head(DC_catch.10)
```

# Marina area 10 catch data
```{r catch two, cache=TRUE}
DC_catch.11.13 <- read.csv("../Data/2007.2017_DC_catch.csv", header = TRUE)
```

```{r table 11.13, echo=FALSE, cache=TRUE}
head(DC_catch.11.13)
```

# Stitching tables
```{r stitch, cache=TRUE}
DC_catch <- inner_join(DC_catch.10, DC_catch.11.13, by = "year")
```

There is now a single table describing the catch rates in lbs. at marine areas 10, 11, and 13.

```{r, cache=TRUE}
head(DC_catch)
```

---

## Plot of main data

```{r main plot, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(DC_catch) +
  geom_point(mapping = aes(x = year, y = lbs_dungeness.11), col = "blue", 
             shape  = "triangle") +
  geom_smooth(mapping = aes(x = year, y = lbs_dungeness.11), col = "blue") +
  geom_point(mapping = aes(x = year, y = lbs_dungeness.13), col = "red") +
  geom_smooth(mapping = aes(x = year, y = lbs_dungeness.13), col = "red") +
  ylim(0,125000) +
  labs(title = "Pounds of Dungeness Caught per Year", x = "Year", 
       y = "Pounds of Dungeness")
```

---

## Analysis

Looking for trends within the main plot.

# Data subsets
```{r subsets, cache=TRUE}
lin_data.10 <- subset(DC_catch, year %in% c("2013", "2014", "2015", "2016", "2017"))
lin_data.11 <- subset(DC_catch, year %in% c("2014", "2015", "2016", "2017"))
lin_data.13 <- subset(DC_catch, year %in% c("2011", "2012", "2013", "2014", 
                                            "2015", "2016", "2017"))
```

Subsets focus in on a particular part of the data showing a decline.

# Linear models using subsets
```{r linear model, cache=TRUE}
lm_fit.10 <- lm(year ~ lbs_dungeness.10, data=lin_data.10)
summary(lm_fit.10)

lm_fit.11 <- lm(year ~ lbs_dungeness.11, data=lin_data.11)
summary(lm_fit.11)

lm_fit.13 <- lm(year ~ lbs_dungeness.13, data=lin_data.13)
summary(lm_fit.13)
```

There is no significant decrease in catch rate in marine area 10, however there is a significant decrease in areas 11 and 13.

---

## Subset plots

# Marine area 11
```{r subset plot area 11, echo=FALSE, warning=FALSE, message=FALSE}
predicted.11 <- data.frame(catch_pred = predict(lm_fit.11, lin_data.11), 
                           lbs_dungeness.11=lin_data.11$lbs_dungeness.11)

ggplot(lin_data.11) +
  geom_point(mapping = aes( x = year, y = lbs_dungeness.11), 
             col = "blue", shape = "triangle", size = 3) +
  geom_line(data = predicted.11, aes(x=catch_pred, y=lbs_dungeness.11)) +
  labs(title = "Pounds of Dungeness Caught per Year", x = "Year", 
       y = "Pounds of Dungeness")
```

# Marine area 13
```{r subset plot area 13, echo=FALSE, warning=FALSE, message=FALSE}
predicted.13 <- data.frame(catch_pred = predict(lm_fit.13, lin_data.13), 
                           lbs_dungeness.13=lin_data.13$lbs_dungeness.13)

ggplot(lin_data.13) +
  geom_point(mapping = aes( x = year, y = lbs_dungeness.13), 
             col = "red", size = 3) +
  geom_line(data = predicted.13, aes(x=catch_pred, y=lbs_dungeness.13)) +
  labs(title = "Pounds of Dungeness Caught per Year", x = "Year", 
       y = "Pounds of Dungeness")
```

